Synthesis of Linear Functional Programs

### 2021-07-23 article/article.tex
add examples

To showcase the expressiveness of our synthesizer, we present as
examples the input and output program for \emph{reverse}.
Note how we can guide the synthesis using the assertion mechanism to 
supply an example:

@@

\paragraph{Maybe}
input:
\begin{minted}{haskell}
data Maybe a = Nothing | Just a;
data List a = Nil | Cons (a * List a);

synth return :: a -o Maybe a;
synth empty :: Maybe a;
synth bind :: Maybe a -o (a -o Maybe b) -> Maybe b;
synth maybe :: b -> (a -o b) -> Maybe a -o b;
\end{minted}
output:
\begin{minted}{haskell}
return :: forall a . (a -o Maybe a);
return = Just;

@@

>>>>

\noindent \emph{Example.} To showcase the expressiveness of our synthesizer, we present as
examples the input and output program for \emph{reverse}.
Note how we can guide the synthesis using the assertion mechanism to 
supply an example:

@@

\mypara{Maybe}\
\\
Input program:
\begin{minted}{haskell}
data Maybe a = Nothing | Just a;
data List a = Nil | Cons (a * List a);

synth return :: a -o Maybe a;
synth empty :: Maybe a;
synth bind :: Maybe a -o (a -o Maybe b) -> Maybe b;
synth maybe :: b -> (a -o b) -> Maybe a -o b;
\end{minted}
Output program:
\begin{minted}{haskell}
return :: forall a . (a -o Maybe a);
return = Just;

@@

\mypara{List}\
\\
Input program:
\begin{minted}{haskell}
data List a = Nil | Cons (a * List a);
data Maybe a = Nothing | Just a;

synth singleton :: a -o List a;
synth append :: List a -o List a -o List a;
synth map :: (!(a -o b)) -o List a -o List b;
synth foldl :: !(b -o a -o b) -o b -o List a -o b | choose 1;
synth uncons :: List a -o Maybe (a * List a);
synth foldr :: !(a -o b -o b) -o b -o List a -o b;
synth insert :: a -o List a -o List a;
synth concat :: List (List a) -o List a;
\end{minted}
Ouput program:
\begin{minted}{haskell}
singleton :: forall a . (a -o List a);
singleton b = Cons (b, Nil);

append :: forall a . (List a -o (List a -o List a));
append b c = case b of
    Nil -> c
  | Cons d ->
      let e*f = d in
        Cons (e, append f c);

map :: forall a b . (!(a -o b) -o (List a -o List b));
map c d = let !e = c in
  case d of
      Nil -> Nil
    | Cons f ->
        let g*h = f in
          Cons (e g, map (!e) h);

foldl :: forall a b . (!(b -o (a -o b)) -o (b -o (List a -o b)));
foldl c d e = let !f = c in
  case e of
      Nil -> d
    | Cons g ->
        let h*i = g in
          foldl (!f) (f d h) i;

uncons :: forall a . (List a -o Maybe (a * List a));
uncons b = case b of
    Nil -> Nothing
  | Cons c ->
      let d*e = c in Just (d, e);

foldr :: forall a b . (!(a -o (b -o b)) -o (b -o (List a -o b)));
foldr c d e = let !f = c in
  case e of
      Nil -> d
    | Cons g ->
        let h*i = g in
          f h (foldr (!f) d i);

insert :: forall a . (a -o (List a -o List a));
insert b c = case c of
    Nil -> Cons (b, Nil)
  | Cons g ->
      let h*i = g in
        Cons (h, insert b i);

concat :: forall a . (List (List a) -o List a);
concat b = case b of
    Nil -> Nil
  | Cons c ->
      let d*e = c in
        case d of
            Nil -> concat e
          | Cons k ->
              let l*m = k in
                Cons (l, concat (Cons (m, e)));
\end{minted}

\mypara{State} (with a slight optimization that will be added as a control keyword
futurely, that allows bind using runState to terminate in a reasonable time)
\\
Input program:
\begin{minted}{haskell}
data State b a = State (!b -o (a * !b));

synth runState :: State b a -o (!b -o (a * !b));
synth bind :: (State c a -o (a -o State c b) -o State c b) | using (runState);
synth return :: a -o State b a;
synth get :: State a a;
synth put :: !a -o (State a 1);
synth modify :: (!a -o !a) -o State a 1;
synth evalState :: State b a -o !b -o a;
\end{minted}
Output program:
\begin{minted}{haskell}
data State b a = State (!b -o (a * !b));

runState :: forall a b . (State b a -o (!b -o (a * !b)));
runState c !f = case c of
    State e ->
        let h*i = e (!f) in
          let !j = i in (h, (!j));

bind :: forall a b c . (State c a -o ((a -o State c b) -o State c b));
bind bs bt = case bs of
    State bu ->
      State (λbv -> let !bw = bv in
                      let cu*cv = bu (!bw) in
                        let !cw = cv in
                          (let dr*ds = runState (bt cu) (!cw) in
                             let !dt = ds in dr, (!cw)));

return :: forall a b . (a -o State b a);
return c = State (λd -> let !e = d in
               (c, (!e)));

get :: forall a . State a a;
get = State (λb -> let !c = b in
               (c, (!c)));

put :: forall a . (!a -o State a 1);
put b = let !c = b in
  State (λd -> let !e = d in
                 ((), (!e)));

modify :: forall a . ((!a -o !a) -o State a 1);
modify b = State (λc -> let !d = c in
               let !f = b (!d) in ((), (!f)));

evalState :: forall a b . (State b a -o (!b -o a));
evalState c d = case c of
    State e ->
      let !f = d in
        let h*i = e (!f) in
          let !j = i in h;
\end{minted}
